---
title: "EDA"
author: "Melody Jiang"
date: "1/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(reshape2)
library(dplyr)
library(corrplot)
library(mice) # missing data imputation
library(nnet) #multinomial regression models
library(broom)
library(arm) # binned residual plot
```

```{r}
df <- readRDS("Longnecker.rds")
```

```{r}
summary(df)
```

```{r}
df_no_albumin <- df %>% select(-albumin)
```

```{r}
df_no_albumin <- df_no_albumin %>% mutate(early_delivery=cut(df_no_albumin$gestational_age, c(-Inf, 37, Inf), include.lowest = TRUE, , labels=c("early", "normal")))
df_no_albumin <- df_no_albumin %>% mutate(maternal_age_30=cut(df_no_albumin$maternal_age, c(-Inf, 30, Inf), include.lowest = FALSE, , labels=c("30_and_below", "above_30")))
df_no_albumin_imputed <- mice(df_no_albumin, m=1, maxit = 50, method = 'pmm', seed = 500)
df_no_albumin_imputed <- complete(df_no_albumin_imputed, 1)
```

```{r}
df_temp <- df_no_albumin %>% select(dde:triglycerides, cholesterol) %>% melt()
ggplot(df_temp,aes(x = value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_histogram()
```

```{r}
df_temp <- df_no_albumin %>% select(race:smoking_status, ) %>% melt()
ggplot(df_temp,aes(x = value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_histogram()
```

```{r}
ggplot(df_no_albumin, aes(x = early_delivery, y = dde, group = early_delivery)) +
  geom_boxplot(aes(fill=early_delivery))
```

```{r}
df_temp <- df_no_albumin %>% select(pcb_028:pcb_203, early_delivery) %>% melt()
ggplot(df_temp, aes(x=early_delivery, y=value, group=early_delivery)) + 
  geom_boxplot(aes(fill=early_delivery)) +
  facet_grid(. ~ variable)
```

```{r}
ggplot(df_no_albumin, aes(x = early_delivery, y = triglycerides, group = early_delivery)) +
  geom_boxplot(aes(fill=early_delivery))
```

```{r}
ggplot(data = df_no_albumin, aes(x = race, fill = df_no_albumin$early_delivery)) + 
    geom_bar()
```

```{r}
df_temp <- as.data.frame(table(df_no_albumin$race, df_no_albumin$early_delivery))
ggplot(df_temp, aes(x = Var1, y = Freq, fill = Var2)) + 
    geom_bar(position = "fill",stat = "identity") +
    # or:
    # geom_bar(position = position_fill(), stat = "identity") 
    scale_y_continuous(labels = scales::percent_format())
```

```{r}
table(df_no_albumin$race, df_no_albumin$early_delivery)
```

```{r}
slices <- c(160, 872) 
lbls <- c("early", "normal")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices,labels = lbls, col=rainbow(length(lbls)),
   main="Early deliveries in white")
```

```{r}
slices <- c(335, 889) 
lbls <- c("early", "normal")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices,labels = lbls, col=rainbow(length(lbls)),
   main="Early deliveries in black")
```

```{r}
slices <- c(27, 97) 
lbls <- c("early", "normal")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices,labels = lbls, col=rainbow(length(lbls)),
   main="Early deliveries in other races")
```

```{r}
ggplot(data = df_no_albumin, aes(x = score_education, fill = early_delivery)) + 
    geom_bar()
```

```{r}
df_temp <- as.data.frame(table(df_no_albumin$score_education, df_no_albumin$early_delivery))
ggplot(df_temp, aes(x = Var1, y = Freq, fill = Var2)) + 
  geom_bar(position = "fill",stat = "identity") +
  # or:
  # geom_bar(position = position_fill(), stat = "identity") 
  scale_y_continuous(labels = scales::percent_format()) +
  ggtitle("Percentage of early delivery versus education level")
```

```{r}
ggplot(data = df_no_albumin, aes(x = score_income, fill = early_delivery)) + 
    geom_bar()
df_temp <- as.data.frame(table(df_no_albumin$score_income, df_no_albumin$early_delivery))
ggplot(df_temp, aes(x = Var1, y = Freq, fill = Var2)) + 
  geom_bar(position = "fill",stat = "identity") +
  # or:
  # geom_bar(position = position_fill(), stat = "identity") 
  scale_y_continuous(labels = scales::percent_format()) +
  ggtitle("Percentage of early delivery versus income score")
```

```{r}
ggplot(data = df_no_albumin, aes(x = score_occupation, fill = early_delivery)) + 
    geom_bar()
df_temp <- as.data.frame(table(df_no_albumin$score_occupation, df_no_albumin$early_delivery))
ggplot(df_temp, aes(x = Var1, y = Freq, fill = Var2)) + 
  geom_bar(position = "fill",stat = "identity") +
  # or:
  # geom_bar(position = position_fill(), stat = "identity") 
  scale_y_continuous(labels = scales::percent_format()) +
  ggtitle("Percentage of early delivery versus occupation")
```

```{r}
ggplot(data = df_no_albumin, aes(x = maternal_age, fill = early_delivery)) + 
    geom_bar()
```

```{r}
table(df_no_albumin$maternal_age_30, df_no_albumin$early_delivery)
```

```{r}
slices <- c(428, 1551) 
lbls <- c("early", "normal")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices,labels = lbls, col=rainbow(length(lbls)),
   main="Early deliveries in mothers aged 30 and below")
slices <- c(94, 307) 
lbls <- c("early", "normal")
pct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices,labels = lbls, col=rainbow(length(lbls)),
   main="Early deliveries in mothers above 30")
```

```{r}
ggplot(data = df_no_albumin, aes(x = smoking_status, fill = early_delivery)) + 
    geom_bar()
df_temp <- as.data.frame(table(df_no_albumin$smoking_status, df_no_albumin$early_delivery))
ggplot(df_temp, aes(x = Var1, y = Freq, fill = Var2)) + 
  geom_bar(position = "fill",stat = "identity") +
  # or:
  # geom_bar(position = position_fill(), stat = "identity") 
  scale_y_continuous(labels = scales::percent_format()) +
  ggtitle("Percentage of early delivery versus smoking status")
```

```{r}
ggplot(data = df_no_albumin, aes(x = center, fill = early_delivery)) + 
    geom_bar()
df_temp <- as.data.frame(table(df_no_albumin$center, df_no_albumin$early_delivery))
ggplot(df_temp, aes(x = Var1, y = Freq, fill = Var2)) + 
  geom_bar(position = "fill",stat = "identity") +
  # or:
  # geom_bar(position = position_fill(), stat = "identity") 
  scale_y_continuous(labels = scales::percent_format()) +
  ggtitle("Percentage of early delivery versus center")
```

```{r}
corrplot(cor(df_no_albumin_imputed %>% dplyr::select(dde:triglycerides, cholesterol)))
```


# Observations

Chemicales are very skewed to the right.

There is no visible difference in chemicales as of early delivery.

Multicolinearity between pcb's -  might want to deal with this.


# Initial Analysis

We begin by building a multinomial logistic model that only includes the main eﬀects.

```{r}
model1 <- multinom(data = df_no_albumin_imputed, early_delivery ~ . - maternal_age_30)
```

```{r}
tidy(model1)
```

We see statistically insignificat p-values.

We look at binned residual plot for suggestions of model improvement.

```{r}
# Calculate predicted probabilities
pred.probs <-  predict(model1,type="probs")
resid <- if_else(df_no_albumin_imputed$early_delivery==1, 1, 0) - pred.probs
```

```{r}
hist(pred.probs)
```

```{r}
# binned residual plot for dde 
p_1 <- binnedplot(x = df_no_albumin_imputed$dde ,y = resid, nclass=NULL, 
                  xlab="dde", ylab="Average residual", 
                  main="Binned residual plot: residual vs dde", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
p_2 <- binnedplot(x = df_no_albumin_imputed$pcb_028 ,y = resid, nclass=NULL, 
                  xlab="pcb_028", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_028", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
p_3 <- binnedplot(x = df_no_albumin_imputed$pcb_153 ,y = resid, nclass=NULL, 
                  xlab="pcb_153", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_153", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
p_3 <- binnedplot(x = df_no_albumin_imputed$cholesterol ,y = resid, nclass=NULL, 
                  xlab="cholesterol", ylab="Average residual", 
                  main="Binned residual plot: residual vs cholesterol", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
```


# Dealing with multicolinearity in model
```{r}
df_no_albumin_imputed <- df_no_albumin_imputed %>% mutate(sum_pcb = pcb_074 + pcb_105 + pcb_118 + pcb_153 + pcb_138 + pcb_170 + pcb_180 + pcb_194 + pcb_203)
```

```{r}
corrplot(cor(df_no_albumin_imputed %>% dplyr::select(dde, pcb_028, pcb_052, sum_pcb_1, sum_pcb_2, triglycerides, cholesterol)))
```

```{r}
df <- df %>% mutate(gestational_age_factor = factor(ifelse(gestational_age < 37, "early", "normal")))
df <- df %>% mutate(gestational_age_factor_2 = ifelse(gestational_age < 37, 1, 0))
df_no_albumin_imputed <- cbind(df_no_albumin_imputed, df$gestational_age_factor_2)
```

```{r}
dplyr::rename(df_no_albumin_imputed, df$gestational_age_factor_2 = gestational_age_factor_2)
```

```{r}
model2 <- multinom(data = df_no_albumin_imputed, df$gestational_age_factor_2 ~ dde + pcb_028 + pcb_052 + sum_pcb_1 + sum_pcb_2 + triglycerides + cholesterol + race + score_education + score_income + score_occupation + maternal_age + smoking_status + center)
```

```{r}
tidy(model2)
```

```{r}
# Calculate predicted probabilities
pred.probs <-  predict(model2,type="probs")
resid <- if_else(df_no_albumin_imputed$"df$gestational_age_factor_2"==1, 1, 0) - pred.probs
```

```{r}
# binned residual plot for dde 
p_1 <- binnedplot(x = df_no_albumin_imputed$dde ,y = resid, nclass=NULL, 
                  xlab="dde", ylab="Average residual", 
                  main="Binned residual plot: residual vs dde", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
p_2 <- binnedplot(x = df_no_albumin_imputed$pcb_028 ,y = resid, nclass=NULL, 
                  xlab="pcb_028", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_028", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
p_3 <- binnedplot(x = df_no_albumin_imputed$pcb_052 ,y = resid, nclass=NULL, 
                  xlab="pcb_052", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_052", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
p_4 <- binnedplot(x = df_no_albumin_imputed$sum_pcb_1 ,y = resid, nclass=NULL, 
                  xlab="sum_pcb_1", ylab="Average residual", 
                  main="Binned residual plot: residual vs sum_pcb_1", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
p_5 <- binnedplot(x = df_no_albumin_imputed$sum_pcb_2 ,y = resid, nclass=NULL, 
                  xlab="sum_pcb_2", ylab="Average residual", 
                  main="Binned residual plot: residual vs sum_pcb_2", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
p_6 <- binnedplot(x = df_no_albumin_imputed$triglycerides ,y = resid, nclass=NULL, 
                  xlab="triglycerides", ylab="Average residual", 
                  main="Binned residual plot: residual vs triglycerides", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
p_7 <- binnedplot(x = df_no_albumin_imputed$cholesterol ,y = resid, nclass=NULL, 
                  xlab="cholesterol", ylab="Average residual", 
                  main="Binned residual plot: residual vs cholesterol", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
```



# Quesitons

Multinomial logistic regression what does the "converged" do?

Initial model including only main effects show insignificant p-values. What are some useful steps to take next?

What is the output of a logistic regression here? Why do we use binned residual plot and how do we interpret them? How does binned residual plot work?

Data is quite skewed. Mean imputation good?



