---
title: "Analysis_1"
author: "Melody Jiang"
date: "1/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
set.seed(42)
```

```{r}
library(ggplot2)
library(reshape2)
library(dplyr)
library(corrplot)
library(mice) # missing data imputation
library(nnet) #multinomial regression models
library(broom)
library(arm) # binned residual plot
library(car)
```

```{r}
df <- readRDS("Longnecker.rds")
```

```{r}
df <- df %>% dplyr::select(-albumin) # remove albumin because of too many missing values
df <- df %>% mutate(early_delivery = ifelse(gestational_age < 37, 1, 0))
```

```{r}
summary(df)
```

```{r}
# remove row with na in pcb
df <- df[!is.na(df$pcb_028), ]
```

```{r}
summary(df) # now only covariates have missingness
```

```{r}
# impute missing value

```

```{r}
corrplot::corrplot(cor(df_imputed %>% dplyr::select(-race)))
```

We first examine results including only chemicals.

```{r}
model1 <- glm(early_delivery ~ ., 
              data = df %>% dplyr::select(dde:pcb_203, triglycerides, cholesterol, early_delivery),
              family = "binomial")
```

```{r}
summary(model1)
```

We have seen high correlation between some variables, so p-value here might not be trust-worthy. Since our goal is to exammine associations, it is important to address problems with multicolinearity. We examine variation inflation factors (VIF) for multicolinearity.

```{r}
car::vif(model1)
```

We see that `pcb_105`, `pcb_118`, `pcb_153`, `pcb_138`, `pcb_180` shows moderate to high multicolinearity. `pcb_028`, `pcb_052`, `pcb_074`, `pcb_170`, `pcb_194`, `pcb_203` in pcb's shows acceptable multicolinearity.

Next, we attempt mean centering chemicals, as normalizing variables could remove structural multicolinearity.

```{r}
df_chem <- df %>% dplyr::select(dde:pcb_203, triglycerides, cholesterol)
df_cov <- df %>% dplyr::select(-(dde:pcb_203), -triglycerides, -cholesterol)
df_chem_centered <- cbind(lapply(df_chem, function(x){scale(x, scale = FALSE)}),
      df_cov)
```

We perform analysis and evaluation again.

```{r}
model2 <- glm(early_delivery ~ ., 
              data = df_chem_centered %>% dplyr::select(dde:pcb_203, triglycerides, cholesterol, early_delivery),
              family = "binomial")
summary(model2)
```

```{r}
car::vif(model2)
```

We see that mean-centering does not work. Let's try adding those with moderate to high multicolinearity (VIF >= 5).

```{r}
df <- df %>% mutate(pcb_sum = pcb_105 + pcb_118 + pcb_153 + pcb_138 + pcb_180)
```

```{r}
corrplot::corrplot(cor(df %>% dplyr::select(dde, pcb_028, pcb_052, pcb_074, pcb_170, pcb_194,
                                            pcb_203, pcb_sum, triglycerides, cholesterol, early_delivery)))
```

```{r}
model3 <- glm(early_delivery ~ ., 
              data = df %>% dplyr::select(dde, pcb_028, pcb_052, pcb_074, pcb_170, pcb_194,
                                                        pcb_203, pcb_sum, triglycerides, cholesterol, early_delivery),
              family = "binomial")
summary(model3)
```

```{r}
car::vif(model3)
```

Try other adding ups...

```{r}
df <- df %>% mutate(pcb_sum_2 = pcb_105 + pcb_153 + pcb_138 + pcb_180)
```

```{r}
corrplot::corrplot(cor(df %>% dplyr::select(dde, pcb_028, pcb_052, pcb_074, pcb_170, pcb_194,
                                            pcb_203, pcb_118, pcb_sum_2, triglycerides, cholesterol, early_delivery)))
```

```{r}
model4 <- glm(early_delivery ~ ., 
              data = df %>% dplyr::select(dde, pcb_028, pcb_052, pcb_074, pcb_170, pcb_194,
                                                        pcb_203, pcb_118, pcb_sum_2, triglycerides, cholesterol, early_delivery),
              family = "binomial")
summary(model4)
```

```{r}
confint(model3, 'dde', level=0.95)
```

```{r}
car::vif(model4)
```

We see that vif of model 4 is not as good as model 3.

We temporarily work from model 3 from now. Next, we check residual plots.

```{r}
pred.probs <-  predict(model3, type="response") # Calculate predicted probabilities
resid <- if_else(df$early_delivery==1, 1, 0) - pred.probs # residuals
```

```{r}
# residual plot
plot(pred.probs, resid, main="Residual plot", 
   xlab="estimated probabilities", ylab="residual", pch=19)
abline(a = 0, b = 0)
```

```{r}
binnedplot(x = pred.probs ,y = resid, nclass=NULL, 
            xlab="dde", ylab="Average residual", 
            main="Binned residual plot: residual vs estimated probabilities", 
            cex.pts=0.8, col.pts=1, col.int="gray")
```

```{r}
binnedplot(x = df$dde ,y = resid, nclass=NULL, 
            xlab="dde", ylab="Average residual", 
            main="Binned residual plot: residual vs dde", 
            cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_028 ,y = resid, nclass=NULL, 
                  xlab="pcb_028", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_028", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_052 ,y = resid, nclass=NULL, 
                  xlab="pcb_052", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_052", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_074 ,y = resid, nclass=NULL, 
                  xlab="pcb_074", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_074", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_170 ,y = resid, nclass=NULL, 
                  xlab="pcb_170", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_170", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_194 ,y = resid, nclass=NULL, 
                  xlab="pcb_194", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_194", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_203 ,y = resid, nclass=NULL, 
                  xlab="pcb_203", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_203", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_118 ,y = resid, nclass=NULL, 
                  xlab="pcb_118", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_118", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_sum_2 ,y = resid, nclass=NULL, 
                  xlab="pcb_sum_2", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_sum_2", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$triglycerides ,y = resid, nclass=NULL, 
                  xlab="triglycerides", ylab="Average residual", 
                  main="Binned residual plot: residual vs triglycerides", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$cholesterol ,y = resid, nclass=NULL, 
                  xlab="cholesterol", ylab="Average residual", 
                  main="Binned residual plot: residual vs cholesterol", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
```

```{r}
# Graphing the fitted model
jitter.binary <- function(a, jitt = .05){
  ifelse(a == 0, runif(length(a), 0, jitt), runif(length(a), 1 - jitt, 1))
}
```

```{r}
early.jitter <- jitter.binary (df$early_delivery)
plot (df$dde, early.jitter)
curve (invlogit (coef(model3) [1] + coef(model3) [2]*x), add=TRUE)
```

Interactions between fat chemicales and harmful chemicals. We first mean center our variables.

```{r}
df_chem <- df %>% dplyr::select(dde:pcb_203, triglycerides, cholesterol, pcb_sum)
df_cov <- df %>% dplyr::select(-(dde:pcb_203), -triglycerides, -cholesterol, -pcb_sum, -pcb_sum_2)
df_chem_centered <- cbind(lapply(df_chem, function(x){scale(x, scale = FALSE)}),
      df_cov)
```

```{r}
model5 <- glm(early_delivery ~ . + dde:triglycerides + pcb_028:triglycerides + pcb_052:triglycerides + pcb_074:triglycerides + pcb_170:triglycerides + pcb_194:triglycerides + pcb_203:triglycerides + pcb_sum:triglycerides + dde:cholesterol + pcb_028:cholesterol + pcb_052:cholesterol + pcb_074:cholesterol + pcb_170:cholesterol + pcb_194:cholesterol + pcb_203:cholesterol + pcb_sum:cholesterol, 
              data = df_chem_centered %>% dplyr::select(dde, pcb_028, pcb_052, pcb_074, pcb_170, pcb_194,
                                                        pcb_203, pcb_sum, triglycerides, cholesterol, early_delivery),
              family = "binomial")
summary(model5)
```

```{r}
car::vif(model5)
```

```{r}
pred.probs_5 <-  predict(model5, type="response") # Calculate predicted probabilities
resid_5 <- if_else(df$early_delivery==1, 1, 0) - pred.probs # residuals
```

```{r}
# residual plot
plot(pred.probs_5, resid_5, main="Residual plot", 
   xlab="estimated probabilities", ylab="residual", pch=19)
abline(a = 0, b = 0)
```

```{r}
binnedplot(x = pred.probs_5 ,y = resid_5, nclass=NULL, 
            xlab="dde", ylab="Average residual", 
            main="Binned residual plot: residual vs estimated probabilities", 
            cex.pts=0.8, col.pts=1, col.int="gray")
```

```{r}
binnedplot(x = df$dde ,y = resid_5, nclass=NULL, 
            xlab="dde", ylab="Average residual", 
            main="Binned residual plot: residual vs dde", 
            cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_028 ,y = resid_5, nclass=NULL, 
                  xlab="pcb_028", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_028", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_052 ,y = resid_5, nclass=NULL, 
                  xlab="pcb_052", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_052", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_074 ,y = resid_5, nclass=NULL, 
                  xlab="pcb_074", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_074", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_170 ,y = resid_5, nclass=NULL, 
                  xlab="pcb_170", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_170", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_194 ,y = resid_5, nclass=NULL, 
                  xlab="pcb_194", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_194", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_203 ,y = resid_5, nclass=NULL, 
                  xlab="pcb_203", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_203", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_118 ,y = resid_5, nclass=NULL, 
                  xlab="pcb_118", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_118", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$pcb_sum ,y = resid_5, nclass=NULL, 
                  xlab="pcb_sum_2", ylab="Average residual", 
                  main="Binned residual plot: residual vs pcb_sum_2", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$triglycerides ,y = resid_5, nclass=NULL, 
                  xlab="triglycerides", ylab="Average residual", 
                  main="Binned residual plot: residual vs triglycerides", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
binnedplot(x = df$cholesterol ,y = resid_5, nclass=NULL, 
                  xlab="cholesterol", ylab="Average residual", 
                  main="Binned residual plot: residual vs cholesterol", 
                  cex.pts=0.8, col.pts=1, col.int="gray")
```

```{r}
anova(model3, model5) # doesn't tell me significance?
```

```{r}
# Jiyi mentioned adding all pcb's together. Let's try this.
df_chem_centered <- df_chem_centered %>% mutate(pcb_sum_all = rowSums(df_chem_centered %>% dplyr::select(pcb_028:pcb_203)))
```

```{r}
model6 <- glm(early_delivery ~ ., 
              data = df_chem_centered %>% dplyr::select(dde, pcb_sum_all, triglycerides, cholesterol, early_delivery),
              family = "binomial")
summary(model6)
```

```{r}
car::vif(model6)
```

```{r}
# Add in interactions
model7 <- glm(early_delivery ~ . + dde:triglycerides + dde:cholesterol + pcb_sum_all:triglycerides + pcb_sum_all:cholesterol, 
              data = df_chem_centered %>% dplyr::select(dde, pcb_sum_all, triglycerides, cholesterol, early_delivery),
              family = "binomial")
summary(model7)
```

```{r}
car::vif(model7)
```

```{r}
df_chem_centered <- df_chem_centered %>%
  mutate(race = relevel(race, ref = "white"))
```

```{r}
model8 <- glm(early_delivery ~ . + dde:triglycerides + dde:cholesterol + pcb_sum_all:triglycerides + pcb_sum_all:cholesterol, 
              data = df_chem_centered %>% dplyr::select(dde, pcb_sum_all, triglycerides, cholesterol, early_delivery, race:smoking_status, center),
              family = "binomial")
summary(model8)
```

```{r}
car::vif(model8)
```

# TO DO

make scores ordered categorical data?
make smoking status factors
make center factors

interpreting vif
justification of normalizing variables can help with multicolinearity.

dummy variables for factors?

Class bias?

Potential of using factor models?

Interpretation of residual plot

Scale

Variable selection?


# Questions

How to deal with missingness in scores, ordered discrete data.

In what situations would we want to normalize data and is it always a good practice?

How do we decide what variables to add together though?

What unit is a significant unit?

Dealing with LOD - 



