\documentclass{beamer}

\usepackage[utf8]{inputenc}

\title{Case Study 1-Group 1}
\author{Melody Jiang, Irene Ji, Keru Wu}
\institute{Department of Statistical Science, Duke University}
\date{01/21/2019}

\begin{document}

\frame{\titlepage}



%%%%%%%%%% Section1: Introduction %%%%%%%%%%%


\begin{frame}
\frametitle{Introduction}

\begin{itemize}
\item Data: A study by Longnecker et al. (2001), comprised of 2380 observations of pregnant women.
  
\item Goal: Assess how DDE and PCBs relate to risk of premature delivery.

\end{itemize}

\end{frame}






%%%%%%%%%% Section2: Materials & Methods %%%%%%%%%%%

\begin{frame}
\frametitle{Data Preprocessing}
\begin{itemize}

\item Preterm birth: Gestational Age $\leq$ 36.
\pause
\item Standardize continuous variables.
\pause
\item Missing data: Multivariate Imputations by Chained Equations (MICE package in R) for covariates. Only one missing in continuous, deleted.
\pause
\item Limit of Detection (LOD): Exists in some PCBs. All LODs are negligible compared to data scale (e.g. 0.01 compared to 0.3)

\end{itemize}
\end{frame}




\begin{frame}
\frametitle{Collinearity \& Dimensionality Reduction}

\begin{itemize}

\item There are 11 types of PCBs, some of which have high correlation and may be redundant.
\pause
\item Possible approaches: Simple sum, PCA, Factor Analysis.
\pause
\item Check Multicollinearity: Variation Inflation Factor (VIF).

\end{itemize}
\end{frame}







\begin{frame}
\frametitle{Model}


\begin{itemize}
\item Linear regression or logistic regression?
\pause
\item Too simple $\&$ Can't fit nonlinear trend.
\pause
\item Domain knowledge?
\pause
\item Chemicals have no effect when concentration is lower than a bound.
\pause
\item Chemicals have constant effect when concentration is higher than a bound.
\pause
\item Nonlinear Model
\end{itemize}
\end{frame}











\begin{frame}
\frametitle{Model}


\begin{itemize}
\item Generalized Additive Model (GAM)
$$g(Y_i) = \beta_0 + \sum_{j=1}^m f_i(x_{ij}) + \sum_{k=1}^l \beta_{k}z_{ik}$$

\item Choice of $g$: probit or logit.
\item $x_{.j}$s include DDE, PCBs, maternal age, etc.
\item $z_{.k}$s include categorical variables and some confounding variables.

\end{itemize}
\end{frame}




\begin{frame}
\frametitle{Model}

\begin{itemize}

\item Frequentist approach may overestimate uncertainty.
\item Frequentist GAM may produce a non-significant p-value.
\pause
\item Bayesian Generalized Additive Model
$$g(Y_i) = \beta_0 + \sum_{j=1}^m f_j(x_{ij}) + \sum_{k=1}^l \beta_{k}z_{ik}$$

\item Adds priors on the common regression coefficients, priors on the standard deviations of the smooth terms.


\end{itemize}
\end{frame}




\begin{frame}
\frametitle{Discussion}

\begin{itemize}

\item Deal with different centers
\pause
\item Approach 1: Bayesian Hierarchical Model
\pause
\item Approach 2: Mixed Effect / Random Effect Model
\pause
\item Generalized Additive Mixed Model (GAMM)
\item Bayesian GAMM 


\end{itemize}
\end{frame}


\begin{frame}
\frametitle{Discussion}

\begin{itemize}


\item Specialized prior may give narrower credible intervals.
\pause
\item Including Interactions: Bayesian Factor Analysis (Ferrari, F. and Dunson, D.B. 2019)



\end{itemize}
\end{frame}




















\end{document}
