---
title: "Appendix for STA723 Case Study - Group 1"
author: "Melody Jiang, Irene Ji, Keru Wu"
date: "1/22/2020"
output:
  pdf_document
toc: true
number_sections: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This appendix mainly contains codes and additional outputs.


# 1 EDA

## 1.1 

```{r}
dat = readRDS("Longnecker.rds")
dat$center = factor(dat$center)
dat$smoking_status = factor(dat$smoking_status)

dat = dat[-1861,]
library(mice)
dat = dat[,!names(dat) %in% c('albumin')]
imp = mice(dat)
dat = complete(imp)

dat$ind_37 = dat$gestational_age < 37
dat[,c(1:12,13,15,16,17,18,20)] = scale(dat[,c(1:12,13,15,16,17,18,20)])

```

# 2 PCA

## 2.1 PCA results

```{r}
## PCA

pca = princomp(dat[,2:12])
pcb = as.matrix(dat[,2:12])
pcb_pc = pcb %*% pca$loadings
print(pca$loadings)
summary(pca)

dat$PC1 = pcb_pc[,1]
dat$PC2 = pcb_pc[,2]
dat$PC3 = pcb_pc[,3]
dat$PC4 = pcb_pc[,4]


```

# 3 GAM model

# 4 Bayesian GAM

Bayesian Generalized Additive Model
$$g(Y_i) = \beta_0 + \sum_{j=1}^m f_j(x_{ij}) + \sum_{k=1}^l \beta_{k}z_{ik}$$

We add priors on the common regression coefficients, priors on the standard deviations of the smooth terms. The priors are set by default in $rstanarm$ package, which is a weak informative normal prior. 


## 4.1 Model results

```{r}

library(rstanarm)

b_ga = stan_gamm4(ind_37 ~ s(dde) + s(PC1) + s(PC2)+
                    s(triglycerides) + race + score_education + score_income + score_occupation +
                    s(maternal_age) + smoking_status + s(cholesterol) + center,
                   family = binomial(link = 'logit'), data = dat,
                  chain = 1, iter=1000)

plot_nonlinear(b_ga)

```


## 4.2 Model check

```{r}

library(arm)

pred.probs_gam <-  predict(b_ga, dat, type = 'response') # Calculate predicted probabilities
resid_gam <- residuals(b_ga) # residuals

binnedplot(x = pred.probs_gam ,y = resid_gam, nclass=NULL, 
            xlab="Predicted value", ylab="Average residual", 
            main="Binned residual plot: residual vs estimated probabilities for gam", 
            cex.pts=0.8, col.pts=1, col.int="gray")

```
